---
tags: [r]
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
set.seed(47)
```
# pipecleaner

[![Travis build status](https://travis-ci.org/alistaire47/pipecleaner.svg?branch=master)](https://travis-ci.org/alistaire47/pipecleaner)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/agx2ht7cpwrnrk0k?svg=true)](https://ci.appveyor.com/project/alistaire47/pipecleaner)
[![Coverage status](https://codecov.io/gh/alistaire47/pipecleaner/branch/master/graph/badge.svg)](https://codecov.io/github/alistaire47/pipecleaner?branch=master)
[![CRAN status](https://www.r-pkg.org/badges/version/pipecleaner)](https://cran.r-project.org/package=pipecleaner)
[![lifecycle](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)

pipecleaner is a utility R package to debug pipelines using the magrittr `%>%` 
pipe. Its `debug_pipeline` launches the debugging browser on the input pipeline 
in a form that allows the user to step through the successive calls of the 
pipeline, examining the output of each successive element.

## Installation

pipecleaner is not currently on CRAN, but can be installed with

``` r
# install.packages("devtools")
devtools::install_github("alistaire47/pipecleaner")
```

## Debugging pipelines

To debug a pipeline, call `debug_pipeline` on the raw code or a character 
vector of code. If no input is supplied and it is called 
from [RStudio](https://www.rstudio.com/products/RStudio/), it will use whatever 
code is highlighed in the source editor as input.

`debug_pipeline` can also be called via 
an [RStudio add-in](https://rstudio.github.io/rstudioaddins/) by highlighting 
the pipeline to debug and then selecting "Debug pipeline in browser" from the 
"Addins" menu.

Once called, `debug_pipeline` will reassemble the pipeline into a function that 
can be debugged in the browser and call the debugger. Each line adds another 
call from the pipeline and prints and the output so the user can see the status 
of the data passed through the pipeline by stepping through the function. 

The data is also stored to a variable called `dot[N]` in each line, 
where `[N]` is the index of the call, making it easy to compare input and 
output data of a step in the pipeline and try out new code formulations in the 
console.

All together, it looks like this:

```{r debug_pipeline}
library(magrittr)
library(pipecleaner)

debug_pipeline(
    x = 1:5 %>% 
        rev %>% 
        {. * 2} %>% 
        sample(replace = TRUE)
)
```

## Bursting pipes

Occasionally it is necessary to restructure code from a piped to an unpiped 
form. @hrbrmstr dubbed this process "pipe bursting": 

<blockquote class="twitter-tweet" data-partner="tweetdeck"><p lang="en" dir="ltr">coined two new phrases whilst making that r pkg script thing:<br><br>- &quot;declawing&quot; == removing purrr as a dependency<br>- &quot;pipe bursting&quot; == moving from clean %&gt;% to sequential &lt;- statements</p>&mdash; boB Rudis (@hrbrmstr) <a href="https://twitter.com/hrbrmstr/status/1019700751268970502?ref_src=twsrc%5Etfw">July 18, 2018</a><script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script></blockquote>


Now `burst_pipes` makes this sort of restructuring simple:

```{r burst_pipes}
burst_pipes(
    x = 1:5 %>% 
        rev %>% 
        {. * 2} %>% 
        .[3] %>% 
        rnorm(1, ., sd = ./10)
  )
```

More specific names can be specified as a character vector:
```{r}
burst_pipes(
    x <- 1:5 %>% 
        rev %>% 
        {. * 2} %>% 
        .[3] %>% 
        rnorm(1, ., sd = ./10),
    names = c("reversed", "doubled", "third", "x")
)
```

`burst_pipes` can also be called via a pair of RStudio add-ins, which replace 
the highlighted code with its restructured form. The "Burst pipes" add-in 
creates names; the "Burst pipes and set names" add-in allows custom names to be 
set.

## Limitations

pipecleaner should successfully debug most pipelines. However, due to its 
structure, it does have known limitations:

- Only the `%>%` pipe is handled, not more exotic pipes like `%$%`. For the 
    moment, this is unlikely to change absent significant demand.
- Nested pipelines—e.g. piping within an anonymous function in 
    `purrr::map`—are ignored; the whole call is treated as one step. 
    
## Related

- [ViewPipeSteps](https://github.com/daranzolin/ViewPipeSteps) is a similar 
    project that calls `View()` after each step in the pipeline.
- magrittr itself contains `debug_pipe`, which is a wrapper around `browser` 
    that returns its input, allowing it to be inserted within a pipeline.
